All Your Graphs Belong to Us

### 2012-07-17 doc/asplos13/grappa.tex
Removing trivial change
grappa system diagram
initial workstealing details + impl

\section{Grappa} \label{sec:grappa}
[trivial change]
In this section we describe the Grappa runtime system (Figure~\ref{fig:grappa}).  Grappa is a software runtime system that one adds to a commodity HPC cluster.  It provides the following software abstractions to applications: \textbf{memory}, a global distributed shared memory and \textbf{tasking}, co-operative multithreading with automatic load balancing.  Under the hood a third component, \textbf{communication}, a network packet scheduler/aggregator is key to achieving high performance, but largely invisible to applications.  We will now describe each of these components.

@@

Grappa context switches between tasks non-preemptively.  Most calls into the Grappa library, however, do initiate a context switch, and tasks can explicitly yield control.  Grappa switches between tasks entirely within usermode, and in our experiments we have measured the context switch time to be only \checkme{40ns}.  This time is kept small because it is done entirely in usermode and only registers specified as caller saved by the compiler need be saved/restored.  Tasks become bound to a thread stack when no other task is runnable, or Grappa determines it is profitable to do so.  \TODO{How does Grappa determine this?}.

Tasks that are not bound to a stack are steelable by the work steeling system within Grappa.  This load balances work across the collection of nodes executing an application built on Grappa.  \TODO{Talk about the work steeling algorithm that we use}.

>>>>

\section{Grappa} \label{sec:grappa}

\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.5\columnwidth]{figs/grappa-crappy}
\begin{minipage}{0.95\columnwidth}
  \caption{\label{fig:grappa} Grappa runtime system \vspace{-4ex}}
\end{minipage}
\vspace{-3ex}
\end{center}
\end{figure}

In this section we describe the Grappa runtime system (Figure~\ref{fig:grappa}).  Grappa is a software runtime system that one adds to a commodity HPC cluster.  It provides the following software abstractions to applications: \textbf{memory}, a global distributed shared memory and \textbf{tasking}, co-operative multithreading with automatic load balancing.  Under the hood a third component, \textbf{communication}, a network packet scheduler/aggregator is key to achieving high performance, but largely invisible to applications.  We will now describe each of these components.

@@

Grappa context switches between tasks non-preemptively. \TODO{too
    vague: Most calls
    into the Grappa library, however, do initiate a context switch}, and
tasks can explicitly yield control.  Grappa switches between tasks
entirely within usermode, and in our experiments we have measured the
context switch time to be only \checkme{40ns}.  This time is kept
small because it is done entirely in usermode and only registers
specified as caller saved by the compiler need be saved/restored.
Tasks become bound to a thread stack when no other task is runnable,
or Grappa determines it is profitable to do so.  \TODO{How does Grappa
    determine this? BrandonM: right now we don't actually spawn beyond
the starting number of stacks, so Grappa only bounds when there exist
idle worker and unbound task}.

\TODO{some of this first sentence might be in the part where we discuss our high
    level solutions to irregular dynamic parallelism}
The combination of running tasks on a large non-uniform system and
dynamic parallelism creates the potential for an imbalance of tasks
across the system. Some form of dynamic load balancing is necessary.
Grappa employs work stealing: if a \checkme{grappa-process} determines it
does not have enough work, it takes some from a \checkme{grappa-process}
that has public, unstarted tasks.

### 2012-07-17 doc/asplos13/implementation.tex
initial workstealing details + impl

Tasks are not allocated any execution resources until the scheduler
decides to run them; when this occurs, tasks are matched with {\em
  workers}. Each worker simply a collection of status bits and a
stack, allocated at a particular core. Workers

@@

Whenever a task yields or suspends, the scheduler makes a decision
about what to do next. First, it determines if it is time to service
the communication resources should be serviced. This is done on a
periodic basis. Second, it determines if any workers with running
tasks are ready to execute; if so, one is scheduled. Finally, if there
are no workers ready to run, but there are tasks waiting to be matched
with workers, an idle worker is woken (or a new worker is spawned),
matched with a task, and scheduled.

@@

\subsubsection{Work stealing}

\TODO{BrandonM or Jacob}

@@

>>>>

Tasks are not allocated any execution resources until the scheduler
decides to run them; when this occurs, tasks are matched with {\em
  workers}. Each worker is simply a collection of status bits and a
stack, allocated at a particular core. Workers

@@

Whenever a task yields or suspends, the scheduler makes a decision
about what to do next. First, it determines if
the communication resources should be serviced. This is done on a
periodic basis. Second, it determines if any workers with running
tasks are ready to execute; if so, one is scheduled. Finally, if there
are no workers ready to run, but there are tasks waiting to be matched
with workers, an idle worker is woken (or a new worker is spawned),
matched with a task, and scheduled.

@@

\subsubsection{Work stealing} \label{subsec:implementation-worksteal}
When a \TODO{grappa process} runs
out of work, it becomes a ``thief'' and asks a random other
\TODO{grappa process} (``victim'') for work. If the victim has tasks
to spare in its public queue, they will be transferred to the thief's public queue.

Most study of work stealing assumes that a processor has one worker
thread and so it only steals when utilization would go to zero. 
%that is,when there are no more local tasks available. 
Grappa, on the other hand, relies on having many worker threads per core. The
core would be fully-utilized if there is always something on the ready
queue. Even if there are many active tasks, if they are all suspended
for long-latency network requests, then the core is underutilized.

choices we've made
-random victims
-steal from FIFO side of queue
-steal when at the moment readyQ empty and task queues empty
-steal min(chunk\_size, half)

Recent work has explored scalable work stealing in terms of efficient
termination detection. 
\TODO{we should reference or put our global task joiner here}

@@

\TODO{Jacob}
